---
title: Citations (Microsoft)
author: 
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)

# Parameters
file_citations <- here::here("data/microsoft/ms_citations.rds")
file_papers <- here::here("data/arxiv/papers.rds")
#===============================================================================

# Code
citations <- read_rds(file_citations)
papers <- read_rds(file_papers)
```

### How many papers were successfully accessed using the Microsoft API?


```{r}
citations %>% 
  count(accessed = !is.na(id_ms), sort = TRUE) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()
```

95% were successfully accessed. 

### How many papers have more than 0 citations?

```{r}
citations %>% 
  count(cited = citations > 0, sort = TRUE) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()
```

### Distribution of citations

```{r}
citations %>% 
  summarize_at(
    vars(citations), 
    list(min = min, max = max, median = median), na.rm = TRUE
  )
```

The median number of citations is 1.

```{r}
quantile(citations$citations, na.rm = TRUE)
```

For papers with at least 1 citation.

```{r}
citations %>% 
  drop_na(citations) %>% 
  filter(citations < quantile(citations, probs = .95)) %>% 
  ggplot(aes(citations)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(
    breaks = scales::breaks_width(width = 2), 
    expand = c(0, 1)
  ) +
  labs(
    title = "Paper citations, excluding the top 5% of papers",
    subtitle = "39% have 0 citations"
  )
```

```{r}
citations %>% 
  drop_na(citations) %>% 
  count(citations, sort = TRUE) %>% 
  mutate(prop = round(n / sum(n) * 100, 3))
```

### How many papers have DOIs (from Microsoft API)?

```{r}
citations %>% 
  count(has_doi = !is.na(doi), sort = TRUE) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()
```

Only 28% have DOIs. 

Combined with the arXiv data, how many papers have DOIs?

```{r}
combined_doi <-
  citations %>% 
  select(id_arxiv, doi_ms = doi) %>% 
  left_join(
    papers %>% select(id, title, doi_arxiv = doi), 
    by = c("id_arxiv" = "id")
  ) %>% 
  mutate_at(vars(contains("doi")), str_to_lower) 
```

Are the DOIs the same for the papers that have both?

```{r}
combined_doi %>% 
  drop_na(starts_with("doi")) %>% 
  count(same = doi_arxiv == doi_ms, sort = TRUE)
```

Most are the same.

How many papers have at least one DOI listed?

```{r}
combined_doi %>% 
  mutate(doi = coalesce(doi_ms, doi_arxiv)) %>% 
  count(has_doi = !is.na(doi), sort = TRUE) %>% 
  mutate(prop = n / sum(n)) %>% 
  ungroup()
```

Only 30%.

### Do all accessed papers have citation data?

```{r}
citations %>% 
  filter(!is.na(id_ms), is.na(citations))
```

There are no papers with a Microsoft ID and no citation data.
