---
title: "SOTA"
author: "Sara Altman"
date: "3/11/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)

# Parameters
file_sota <- here::here("data/sota.rds")

#===============================================================================

sota <- read_rds(file_sota)
```

## Understanding benchmarks

Each task has one benchmark ID.

```{r}
sota %>% 
  group_by(task) %>% 
  summarize(num_benchmarks = n_distinct(benchmark_id, na.rm = TRUE)) %>% 
  count(num_benchmarks)
```

The benchmark IDs are unique identifiers of the tasks. 

```{r}
n_distinct(sota$benchmark_id) == n_distinct(sota$task)
```

Some benchmarks are associated with multiple metrics.

```{r}
sota %>% 
  group_by(benchmark_id, task) %>% 
  summarize(num_metrics = n_distinct(metric_name, na.rm = TRUE)) %>% 
  ungroup() %>% 
  arrange(desc(num_metrics))
```

Does each task belong to a distinct category set?

```{r}
sota %>% 
  select(index, task, categories) %>% 
  mutate(categories = map_chr(categories, ~ sort(.) %>% str_c(collapse = ", "))) %>% 
  group_by(task) %>% 
  summarize(
    distinct_category_groups = n_distinct(categories, na.rm = TRUE), 
    sort = TRUE
  ) %>% 
  pull(sort) %>% 
  all()
```

Yes, each task is mapped to one category set. 

## Categories

```{r}
sota %>% 
  unnest(cols = categories) %>% 
  count(categories) %>% 
  mutate(categories = fct_reorder(categories, n)) %>% 
  ggplot(aes(n, categories)) +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(width = 50)) +
  labs(
    title = "Number of SOTA measurements by category",
    subtitle = "NLP and computer vision have the most data"
  )
```

```{r}
sota %>% 
  unnest(cols = categories) %>% 
  distinct(categories, task) %>% 
  count(categories) %>% 
  mutate(categories = fct_reorder(categories, n)) %>%
  ggplot(aes(n, categories)) +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(width = 50)) +
  labs(
    title = "Number of distinct tasks by category",
    subtitle = "Computer vision and NLP have the most tasks"
  )
```

## Missing metrics

```{r}
sota %>% 
  mutate_at(vars(metric_name, metric_result), list(missing = is.na)) %>% 
  count(metric_name_missing, metric_result_missing, sort = TRUE) %>% 
  mutate(prop = n / sum(n))
```

33% are missing a metric name and a metric value. There are 7 rows missing a metric value, but not a metric name. These are rows where I was not able to turn the metric value into a double (e.g., it was a word). 

Missing metrics by category

```{r}
missing_metrics <-
  sota %>% 
  unnest(cols = categories) %>% 
  group_by(categories) %>% 
  summarize(
    prop_missing = sum(is.na(metric_result)) / n(),
    n = n()
  ) 

missing_metrics %>% 
  arrange(desc(prop_missing))
```

The Robots category is missing 100% of its metrics.

```{r}
missing_metrics %>% 
  mutate(categories = fct_reorder(categories, prop_missing)) %>% 
  ggplot(aes(prop_missing, categories, size = n)) +
  geom_vline(xintercept = 0.5, color = "white", size = 1.5) +
  geom_point() +
  scale_x_continuous(
    breaks = scales::breaks_width(width = .1), 
    limits = c(0, 1)
  ) +
  labs(
    title = "Missing metrics across categories",
    subtitle = "5 categories are missing over 50% of metric results",
    size = "Number of\nmeasurements"
  )
```

Computer vision and NLP both have a lot of data and relatively low proportions of missing data, so they are probably going to be the most reliable categories. 

```{r}
observations <-
  sota %>% 
  drop_na(metric_name, metric_result) %>% 
  count(benchmark_id, metric_name, sort = TRUE)

observations %>% 
  count(n) %>% 
  mutate(prop = nn / sum(nn))
```

41% of the (non-missing) task-metric combinations have only 1 data point, which means we won't be able to understand how the metric has changed over time for those task-metric combinations.

```{r}
observations %>% 
  ggplot(aes(n)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Number of observations (metric measurements for a given task)")
```

Many metrics are only measured once.

```{r}
observations_by_category <-
  sota %>% 
  drop_na(metric_result, metric_name) %>% 
  unnest(cols = categories) %>% 
  group_by(categories) %>% 
  count(benchmark_id, metric_name, name = "num_observations") %>% 
  count(num_observations, sort = TRUE, name = "n") %>% 
  ungroup()


observations_by_category %>% 
  group_by(categories) %>% 
  summarize(
    prop_one = n[num_observations == 1] / sum(n),
    n_metrics = sum(n)
  ) %>% 
  mutate(categories = fct_reorder(categories, prop_one, .desc = TRUE)) %>% 
  ggplot(aes(prop_one, categories, size = n_metrics)) +
  geom_point() +
  scale_x_continuous(breaks = scales::breaks_width(0.1)) +
  labs(
    x = "Proportion of metrics with only a single observation",
    title = "Proportion of single observation metrics by category",
    subtitle = "Many metrics are only measured once",
    size = "Number of\ndistinct metrics"
  )
```

100% of the metrics for audio, adversarial, and knowledge base were only measured once, which means we won't be able to get anything meaningful from these categories. 

```{r}
sota %>% 
  drop_na(metric_result, metric_name) %>% 
  count(benchmark_id, metric_name, sort = TRUE) %>% 
  count(n) %>% 
  mutate(prop = nn / sum(nn))
```

40% of all task-metric combinations have only one data point.

```{r}
num_observations %>% 
  uncount(weights = n) %>% 
  group_by(categories) %>% 
  summarize(median_observations = median(num_observations)) %>% 
  mutate(categories = fct_reorder(categories, median_observations)) %>% 
  ggplot(aes(median_observations, categories)) +
  geom_point()
``` 

## Percent change

```{r}
y_limits <- quantile(sota$percent_change, probs = c(0.1, .9), na.rm = TRUE)

sota %>% 
  drop_na(metric_name, metric_result) %>% 
  ggplot(aes(factor(1), percent_change)) +
  geom_boxplot() +
  coord_cartesian(ylim = y_limits) +
  labs(
    title = "Percent change for the middle 80% of the data",
    subtitle = "There are lots of outliers"
  )
```

```{r}
quantile(sota$percent_change, na.rm = TRUE)
```

```{r}
y_limits <- 
    quantile(sota$percent_change, probs = c(0.01, .99), na.rm = TRUE)

sota %>% 
  drop_na(metric_name, metric_result) %>%
  filter(percent_change != 0) %>% 
  unnest(cols = categories) %>% 
  mutate(
    categories = 
      fct_reorder(categories, percent_change, .desc = TRUE)
  ) %>% 
  ggplot(aes(categories, percent_change)) +
  geom_hline(yintercept = 0, size = 1.5, color = "white") +
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = 45)
  ) +
  coord_cartesian(ylim = y_limits)
```

Methodology has the outliers.

## SOTA progress

```{r}
sota %>% 
  drop_na(percent_change, paper_date)   %>% 
  filter(metric_name != "Number of params") %>% 
  filter(categories != "Playing Games") %>% 
  filter(percent_change > 0) %>%
  unnest(categories) %>% 
  ggplot(aes(days_written_after_first_in_benchmark, percent_change, color = categories)) +
  geom_point() 

sota %>% 
  drop_na(percent_change, paper_date)   %>% 
  filter(metric_name != "Number of params") %>% 
  filter(categories != "Playing Games") %>% 
  filter(percent_change > 0) %>%
  unnest(categories) %>% 
  ggplot(aes(days_written_after_first_in_benchmark, percent_change, color = categories)) +
  geom_smooth(se = FALSE)

sota %>% 
  unnest(categories) %>% 
  filter(categories == "Playing Games")
```

```{r}
sota %>% 
  drop_na(percent_change, paper_date)   %>% 
  unnest(categories) %>% 
  filter(
    metric_name != "Number of params",
    categories != "Playing Games",
    percent_change != 0
  ) %>% 
  ggplot(aes(paper_date, percent_change, color = categories)) +
  geom_hline(yintercept = 0) +
  geom_point() 
```

```{r}
tasks <- 
  c(
    "Text Classification",
    "Speech Recognition",
    "Face Detection"
  )

sota %>% 
  filter(task %in% tasks) %>% 
  group_by(task) %>% 
  mutate(n_papers = n()) %>% 
  ungroup() %>% 
  filter(
    n_papers > 1, 
    metric_name != "Number of params"
  ) %>% 
  drop_na(paper_date, percent_change) %>% 
  ggplot(aes(paper_date, percent_change, color = task)) +
  geom_hline(yintercept = 0, color = "white", size = 2) +
  geom_point() 
```

```{r}
sota %>% 
  drop_na(percent_change, paper_date)   %>% 
  unnest(categories) %>% 
  filter(
    metric_name != "Number of params",
    categories == "Natural Language Processing",
    percent_change != 0
  ) %>% 
  ggplot(aes(paper_date, percent_change)) +
  geom_hline(yintercept = 0) +
  geom_point() 
```



```{r}
sota_category("Computer Vision")
```

